A Topic Model of Climate Change Literature

### 2019-07-12 coverletter.tex
before merging in overleaf

\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{fancyhdr}
\pagestyle{fancy}

@@

		\marpartop{\textcolor[HTML]{0067a1}{\footnotesize\textsf{Mercator Research Institute on Global Commons and Climate Change (MCC) gemeinnützige GmbH}\\ \bigskip}%
			\par\textsf{\textbf{Max W. Callaghan}}%
			\par\textsf{PhD Candidate}%
			\par\textsf{Applied Sustainability Science}%
			\par\textsf{Torgauer Str. 12-15}
			\par\textsf{10829 Berlin | Germany}
			\par\textsf{tel +49 (0) 30 338 55 27 - 245}
			\par\textsf{fax +49 (0) 30 338 55 37 - 102}
			\par\textsf{mail callaghan@mcc-berlin.net}
			\par\textsf{web www.mcc-berlin.net}
		}

@@

We believe this letter addresses a crucial challenge affecting the broad climate change community: namely, the field's rapidly expanding literature. This phenomenon - there are more than 45,000 papers published on climate change each year - means understanding the field in its entirety is more difficult than ever. 
This is particularly relevant for the IPCC, given their mandate of providing comprehensive, objective and transparent assessments of he literature.

Moreover, this challenge has also hampered how we have attempted to understand how the IPCC engages with the literature. Such attempts \cite{Bjurström2011, Hulme2010, Victor2015, Corbera2016, Kowarsch2017} have often been based on qualitiative perceptions, or out of date data with insufficient methodological sophistication to contrast patterns in IPCC citations with patterns in the wider literature.

Our study has two innovations which allow us advance the current state of knowledge on the the IPCC and the field it is tasked with assessing. First, we set the complete set of IPCC citations into the context of the wider literature, allowing us to show which disciplines have been over or under-represented.  Contrary to previous claims, the social sciences are over-represented in IPCC reports. This is because the proportion of social science articles among publications cited by the IPCC is higher than the proportion of social science articles among all publications on climate change. 
Second, our use of topic modelling allows us to dig down into the content of articles more or less represented in the IPCC. This provides the basis for our second main conclusion, that literature on technical solutions (such as negative emissions, buildings or cities) is growing fast and, even compared to other fast-growing literatures, has been under-represented in IPCC reports.

These insights have several implications across the community, and we further point out how the map we produce can enhance our understanding of the field and contribute to improving processes involved in the production of global environmental assessments.

>>>>

\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{url}
\usepackage{xifthen}

@@

		\marpartop{\textcolor[HTML]{0067a1}{\footnotesize\textsf{Mercator Research Institute on Global Commons and Climate Change (MCC) gemeinnützige GmbH}\\ \bigskip}%
			\par\textsf{\textbf{Max W. Callaghan}}%
			\par\textsf{PhD Candidate}%
			\par\textsf{Applied Sustainability Science}%
			\par\textsf{Torgauer Str. 12-15}
			\par\textsf{10829 Berlin | Germany}
			\par\textsf{tel +49 (0) 30 338 55 27 - 245}
			\par\textsf{fax +49 (0) 30 338 55 37 - 102}
			\par\textsf{mail callaghan@mcc-berlin.net}
			\par\textsf{web \url{www.mcc-berlin.net}}
		}

@@

``Big literature'' - dealing with a vast and rapidly expanding body of publications - poses new, and largely neglected challenges for climate change assessments and how we understand them. For example, we estimate a total of 350,000 new publications on climate change to emerge (only) in the Web of Science during the Sixth Assessment Cycle of the Intergovernmental Panel on Climate Change spanning from 2015 to 2022. Big data and machine learning methods need to be applied for keeping an overview of developments in the scientific landscape and to analyse how past IPCC assessments have engaged with the avaailable literature. So far, attempts \cite{Bjurström2011, Hulme2010, Victor2015, Corbera2016, Kowarsch2017} have often been based on qualitiative perceptions, or incomplete data with insufficient methodological sophistication.

In this study we use machine-learning to provide a comprehensive, thematic map of the more than 400,000 articles on climate change published in the Web of Science to date. We perceive two innovations that allow us advance the current state of knowledge on the IPCC and the field it is tasked with assessing. First, our thematic map allows us to situate a complete set of IPCC citations into the context of the wider literature and analyse disciplinary representation adequately.  We find that contrary to previous claims, the social sciences are over-represented in IPCC reports.Second, the use of machine learning methods in creating the map allows us to analyse the content of publications more or less represented in the IPCC. We find that literature on technical solutions (such as negative emissions, buildings or cities) is growing fast and, even compared to other fast-growing literatures, has been under-represented in IPCC reports. We point towards a variety of other applications of such a comprehensive, thematic literature map within global environmetal assessments as well as the broad field of research synthesis. Our findings have direct implications for addressing growing demands for more solution-oriented climate change assessments that are also more firmly rooted in the social sciences.

### 2019-07-12 main.tex
before merging in overleaf

\documentclass{article}
\usepackage[a4paper, total={180mm, 260mm}]{geometry}
\usepackage{graphicx}
\usepackage{url}
\usepackage{natbib}
\usepackage{todonotes}
\usepackage{booktabs}
\usepackage{lineno}
\usepackage{color}
%\usepackage{auto-pst-pdf}
\usepackage[colaction]{multicol}
\usepackage{caption}
\usepackage{svg}
\usepackage{authblk}
\usepackage{standalone}

@@

% Title
\title{A Topic Model of Climate Change Literature}
\title{Words, words, words: Mapping the Matter of Climate Change Literature}
\title{A Topography of Climate Change Research}
\author[1,2]{Max Callaghan}
\author[1,2]{Jan Minx}
\author[2]{Piers M. Forster}

@@

		\noindent We live in an age of ``Big Literature'' 
		\cite{Nunez-Mir2016, Minx2017l}, where the science of climate change is expanding exponentially \cite{Grieneisen2011, Haunschild2016}. In the five years since the publication of the last IPCC assessment report \cite{IPCC2014c}, 201,606 papers were published in the Web of Science (WoS) (see Table \ref{tab}). This is almost as much as the 204,585 papers published during the first five assessment periods, a period of nearly 30 years. A total of around 350,000 new publications can be expected for the current sixth assessment cycle of the IPCC, based on current growth patterns. Moreover, the literature has also become more diverse. This is reflected in the expansion of the literature's vocabulary - from 2,000 unique words in the first assessment period to 94,746 words so far in the sixth - indicating that the field has incorporated new content. For example, the zika virus, which was mentioned in 182 articles from 2014-2018, had never before been discussed in the titles or abstracts of articles relating to climate change. Yet it has emerged as a topic of high relevance: the incidence of the virus, the outbreak of which in Brazil in 2016 was declared a public health emergency by the WHO, is set to increase under rising global temperatures \cite{Rao2019}. Similar rapid emergence patterns can be seen for [Intended] Nationally Determined Contributions ([I]NDCs) and Sustainable Development Goals (SDGs) in AR6, and Biochar and REDD (The United Nations Collaborative Programme on Reducing Emissions from Deforestation and Forest Degradation in Developing Countries) in AR5, among others.

@@

		Big literature poses at least three challenges for scientific policy advice and science itself: \emph{First}, established procedures in scientific assessments like those conducted by the IPCC fail to address the exploding literature base. For example, the ratio of studies cited in IPCC reports to the number or relevant studies has declined from 60\% to 20\%  \cite{Minx2017l}, posing a rapidly growing risk of selection bias. More generally, the provision of comprehensive, objective, open and transparent assessments of the available scientific literature, as defined in the principles governing IPCC work \cite{IPCC2013}, is no longer possible for authors or author teams by traditional means. Machine reading and learning methods as well as other data science applications are required to enable an understanding of the field of climate change research at scale (REFs). \emph{Second}, evidence synthesis - the enterprise of reviewing the literature based on a formal and systematic set of methods \cite{Chalmers2002} - becomes increasingly important for aggregating and consolidating the rapidly emerging knowledge and enabling scientific assessments to do their job. Yet, traditional methods of evidence synthesis themselves are pushed to their limits by the large amount of scientific publications. Whilst the field of evidence synthesis technology that tries to streamline human tasks through machine learning at the different stages of the review process is still in its infancy \cite{Beller2018}. \emph{Finally}, overwhelming amounts of literature may be a major reasons why assessments of scientific assessments \cite{Bjurström2011, Hulme2010, Victor2015, Corbera2016} do not offer robust quantifications, for example, for claims about the relationship between report citations and the underlying literature. This may substantially affect findings.

		This study uses topic modelling \cite{Blei2010} to map out the vast body of evidence on climate change. Topic modelling is an unsupervised machine-learning technique, where patterns of word co-occurrences in documents are used to learn a set of topics which can be used to describe the corpus. The word topic derives from the Greek word for place (topos), and by \textit{situating} the documents in a reduced-form projection of their thematic content (see Figure \ref{oecd_topic_map}), we create a \textit{topographic map} of the literature on climate change. Such a systematic engagement with the thematic content of the climate science is missing from the literature so far.%To our knowledge, this is the first
		We apply this map in a second step to understand how the IPCC reports have represented the available climate change literature and re-evaluate claims of bias based on a more comprehensive understanding of the available climate science, and enrich the discussion of representation in the literature to include topics as well as disciplines.

@@

		Figure \ref{oecd_topic_map} shows a \emph{thematic} or \emph{topopographic map} of the 400,000 publications on climate change in our dataset with a total number of 140 topics. The number of topics must be defined exogenously, but the results are robust to different specifications (see methods for examples and a thorough explanation of model selection). Using non-negative matrix factoriziation \cite{Lee1999}, the topics are machine-learned from the papers' abstracts (see methods for details).

@@

		The map shown covers a broad range of topics that occur in meaningful clusters. In general, topics related to climate science and impacts are in the West, while solution-oriented topics are in the East. More fine-grained research areas can also be distinguished. For example, publications related to urban infrastructure (buildings, energy, cement, waste) are located in the East (), physical climate impacts such as sea-level rise (), heat waves (), droughts () or crop yield (losses) () are in the South-West and energy systems are in North-East (). There are larger groups of documents at the fringes of the map that relate mainly to one or two specific topics such as biochar, corals, soil organic carbon, clouds and aerosols, diseases, membranes, absorption or adsorption or CO2 storage. Interestingly, scenarios feature centrally in the map, at the interface between different scientific communities. This corresponds to their integrative nature in IPCC reports (REF). This depiction of the thematic structure of the literature, and its ordering and categorising thereby could be useful for individual communities or for climate change assessments \footnote{We include the list of topic-document combinations in the supplementary materials}.

		The disciplinary composition of this research topography indicated by the different colours in Figure \ref{oecd_topic_map} highlights the dominance of natural sciences in climate change research. More than 60\% of the literature is published in natural science journals. Similarly, only in X out of 140 topics do publications from another field make up the majority of documents, in Y cases the contribution of publications in natural science journals make more than 50\%. We calculate disciplinary entropy of topics as a measure of their degree of interdisciplinarity (see Figure \ref{dis-entropy} and methods for details). While research on health, food, or policy comes from a range of disciplines, research on ice and oceans comes almost exclusively from the natural sciences).

@@

		Finally, the topography shows the thematic evolution of the literature, with topics exhibiting distinct patterns of growth. Fast-growing topics in the last three assessment periods have included, among others, coral, risks, adaptation, hydrogen, buildings, CO2 removal, networks and biochar. Biochar is particularly remarkable in that the sizeable literature which emerged in AR5 was completely absent from the climate change literature beforehand. 
		The identification of new topics as they emerge, particularly as these are discovered without prior knowledge of the literature, can help researchers and assesment-makers to keep abreast of a quickly evolving field.

@@

		Figure \ref{oecd_rep}.a shows that the social sciences were indeed under-represented in the third assessment report, but by the fifth assessment report were over-represented. Likewise, other social sciences than economics have become better represented since AR3  (see figure \ref{subfield}f) with social \& economic geography (4.3\% of the literature), political science (1.0\%), and sociology (0.8\%) showing improved representation in AR5 compared to AR3, and social and economic geography, political science, and other social sciences better represented than economics when taking into account the distribution of studies in the literature.

		This challenges what we think we know about the IPCC. The social sciences, by now, are actually the best represented field, with a share in the literature cited by IPCC reports [1.5] times as high as in the literature at large.  On the other hand the Agricultural Sciences and Engineering \& Technology have been consistently under-represented, with [2.5] and [3.5] times the share of studies in the wider literature than in the literature cited by the IPCC in AR5 respectively.

		The topography allows us to delve deeper into the subject matter that receives more or less attention in the IPCC. Figures \ref{oecd_rep}b and \ref{oecd_rep}c plot the representation of the topics shown in the map. Figure \ref{oecd_rep}c shows that topics more commonly cited by IPCC working group I are older and largely better represented in IPCC reports. These topics, for example \textbf{ozone}, \textbf{oceans}, \textbf{clouds}, \textbf{aerosols} and \textbf{sea levels}, are older, predominantly cited by WG I, and well cited by IPCC reports. They make up some of the core topics of the physical science of climate change.

		The topics in the lower right of the graph are the most pertinent to the question of whether the IPCC is well representing knowledge on climate change. These topics are newer and until now have been under-represented in IPCC reports. They are the topics that could be seen as potential gaps in the IPCC's coverage of the science. Because they are new areas of knowledge, they may be highly salient in a periodic assessment process. These topics are primarily in working group III, on mitigation (see methods section [] for a discussion of the categorisation of topics, including CLC, adsorption and hydrogen, which may more properly be of relevance to WGIII).

		The difference between these under-represented new topics and other new topics that are better represented is intriguing. This difference is visible in figure \ref{evolution-map}, where In AR5, the clusters of documents around the \textbf{adsorption}, \textbf{buildings}, and \textbf{biochar} topics contain few IPCC citations, whereas the clusters around \textbf{food}, \textbf{health}, \textbf{adaptation}, and \textbf{GHGs} contain more. As shown in figure \ref{oecd_rep}c, adsorption buildings and biochar are [4, 3.5 and 3.8] times more prevalent in the literature than in IPCC citations, while food is [0.3] times more prevalent in the literature and health and adaptation are [0.1] and [2,5] times more prevalent in IPCC citations respectively. The IPCC, then, has been better at integrating new knowledge from these topics, and in general better at integrating new knowledge from WG II  than WG III topics.

@@

		Recalling policymakers' demands for more solution-oriented assessments \cite{Kowarsch2017}, we could also interpret the topics that are newer and under-represented as ``solutions-relevant''. Many deal with negative emissions, or with mitigation (or to a lesser extent adaptation) options in the transport, buildings and power sectors. All are arguably related to climate solutions. However, while policymakers' demands for solutions-oriented knowledge were rather about policy options, these under-represented new topics deal with more technical solutions and are found rather in technical disciplines within engineering \& technology and the agricultural sciences.

@@

		\subsection*{Machine-learning for climate change assessments}%, evidence synthesis and science of science} % needs to be reviewed
		Keeping track of developments in the scientific literature is becoming increasingly difficult in the age of big literature. This is becoming a central challenge for global environmental assessments such as those conducted by the IPCC \cite{Minx2017l, Kowarsch2017}, as well as for science of science research (\cite{Zeng2017}), the provision of thorough and comprehensive reviews in specific fields, and the efficient allocation of research funding where it is needed. Ultimately it is a challenge for sound, evidence-based scientific policy advice (\cite{White2019}). Systematic maps and evidence gap maps (\cite{James2016}, \cite{McKinnon2015}) provide crucial tools for tracing the literature, but current approaches are constrained by available resources to answering very specific questions, or are limited by a lack of suitable literature.  On the other hand, current scientometric and research weaving approaches (\cite{Nakagawa2019}) can capture a broader landscape of research, but are too unspecific with regard to the thematic content of literature.

		Elsewhere, we show how machine learning can accelerate the provision of traditional maps of evidence with thousands rather than hundreds of data points (\cite{Lamb2018d}). Here we extend this analysis to vast bodies of literature with hundreds of thousands of publications and show how ``topopgraphic maps'' provide an efficient visualisation of these.

		We acknowledge that the climate change topography presented here is not fully comprehensive. The WoS  covers a substantive, yet by no means exhaustive set of publications. Grey literature, some smaller journals and books, not to mention indigenous knowledge \cite{Ford2016b}, are not covered. Similarly, our query (see methods) requires that the keywords, title or abstract of a publication mentions climate change or a synonym or related term. Whilst this is a reasonable system boundary definition, and follows \cite{Grieneisen2011} exactly, we acknowledge the existence of directly relevant studies for climate change research - for example on energy technologies, energy policies or - that may not meet our criteria and remain outside the scope of our analysis. Yet our decisions are driven by data availability and requirements for a clean, climate-changed focused dataset. Given the large size of our sample (400,000+ publications), we argue that our topography provides a reasonable representation of the research landscape on climate change. Furthermore, unless IPCC citation patterns systematically differ in the not included parts of the literature, the main conclusions of this study are unaffected.

		In addition, our analysis is based on a topic model of publication abstracts. Though it is reasonable to assume that abstracts reflect the themes of the papers they describe, this depends on the quality of the abstracts, and thematic features of the underlying texts may not be adequately represented.  In the absence of comprehensive full-text access and remaining, non-trivial challenges in text extraction from PDFs, we argue that abstracts provide a reasonable unit of analysis.

		We apply our map of the climate change literature to understand the representativeness of IPCC citations and argue that answering related questions requires a clear benchmark. We show that applying this benchmark challenges and confirms some prominent findings around disciplinary representation in and solution-orientation of IPCC assessments.

		While proportional representation of all topics and all disciplines may not be desirable, the analysis presented here can help the IPCC to make quantitatively informed decisions about what to represent. The claim that the IPCC needs to incorporate more social science research may indeed be well-founded, but, in light of the fact shown here that the IPCC over-represents the social science literature that is published, this would imply increases in funding for climate change research in the social sciences. While some might argue that engineering or agricultural research on solutions related topics may be too technical and less relevant to IPCC assessments, in light of the size of the literature in these areas relative to that which is cited by the IPCC it would follow that more social science research needs to be carried out that deals with these topics and produces assessment-ready knowledge.

		We highlight that there are a variety of other immediate applications for such comprehensive topic maps and their rich, underlying data. Rapid, comprehensive maps could make a contribution to making climate change assessment processes in various places more efficient. For example, they could inform the scoping of scientific assessment reports to make sure that reports keep abreast of developments in the literature and remain policy relevant. What's more, they could help in the efficient identification of the best subject area experts across regions, sectors and genders (REFs) and directly increase the transparency with regard to the availability and selection of citations. Beyond the organisation of global environmental assessments, such maps can inform research funding decisions, and help individual researchers identi Future research can explore how further methods, and more comprehensive data, can help drive forward these applications.

		While the size and variety of the literature on climate change presents challenges to global environmental assessments, it represents a wealth of knowledge on vital climate solutions. By shedding light on new and under-reported subjects and by providing quantifications of what we think we know about how assessments and their relationship to the literature, computational mapping can help ensure that this knowledge is made use of efficiently.  Topographic maps  can become part of a toolbox which helps the IPCC continue to be able to fulfil its mandate of providing comprehensive, objective and transparent assessments of an expanding literature.

@@

	%\listoffigures
	\linespread{1}
	%\bibliography{Mendeley}
	\begin{thebibliography}{10}

@@

		\bibitem{Nunez-Mir2016}
		Gabriela~C Nunez-Mir, Basil~V Iannone, Bryan~C Pijanowski, Ningning Kong,
		Songlin Fei, and Richard Fitzjohn.
		\newblock {Automated content analysis: addressing the big literature challenge
			in ecology and evolution}.
		\newblock {\em Methods in Ecology and Evolution}, 7(11):1262--1272, 2016.

@@

		\bibitem{Lee1999}
		D~D Lee and H~S Seung.
		\newblock {Learning the parts of objects by non-negative matrix factorization.}
		\newblock {\em Nature}, 401(6755):788--791, 1999.

@@

		\bibitem{Zeng2017}
		An~Zeng, Zhesi Shen, Jianlin Zhou, Jinshan Wu, Ying Fan, Yougui Wang, and
		H.~Eugene Stanley.
		\newblock {The science of science: From the perspective of complex systems}.
		\newblock {\em Physics Reports}, 714-715:1--73, 2017.

		\bibitem{White2019}
		Howard White.
		\newblock {The twenty-first century experimenting society: the four waves of
			the evidence revolution}.
		\newblock {\em Palgrave Communications}, 5(1):1--7, 2019.

		\bibitem{James2016}
		Katy~L. James, Nicola~P. Randall, and Neal~R. Haddaway.
		\newblock {A methodology for systematic mapping in environmental sciences}.
		\newblock {\em Environmental Evidence}, 5(1):1--13, 2016.

		\bibitem{McKinnon2015}
		Madeleine~C. McKinnon, Samantha~H. Cheng, Ruth Garside, Yuta~J. Masuda, and
		Daniel~C. Miller.
		\newblock {Sustainability: Map the evidence}.
		\newblock {\em Nature}, 528(7581):185--187, 2015.

		\bibitem{Nakagawa2019}
		Shinichi Nakagawa, Gihan Samarasinghe, Neal~R. Haddaway, Martin~J. Westgate,
		Rose~E. O'Dea, Daniel~W.A. Noble, and Malgorzata Lagisz.
		\newblock {Research Weaving: Visualizing the Future of Research Synthesis}.
		\newblock {\em Trends in Ecology and Evolution}, 34(3):224--238, 2019.

		\bibitem{Lamb2018d}
		William~F. Lamb, Felix Creutzig, Max~W. Callaghan, and Jan~C. Minx.
		\newblock {Learning about urban climate solutions}.
		\newblock {\em Nature Climate Change}, 9(4):279--287, 2018.

@@

	\end{thebibliography}
	\bibliographystyle{unsrt}

	\include{methods}

>>>>

\documentclass{article}
\usepackage[a4paper, total={180mm, 260mm}]{geometry}
\usepackage{graphicx}
\usepackage{url}
\usepackage{natbib}
\usepackage{todonotes}
\usepackage{booktabs}
\usepackage{lineno}
\usepackage{color}
%\usepackage{auto-pst-pdf}
\usepackage[colaction]{multicol}
\usepackage{caption}
\usepackage{svg}
\usepackage{authblk}
\usepackage{standalone}
\usepackage{xr}
\externaldocument{methods}

@@

% Title
\title{A Topic Model of Climate Change Literature}
\title{Words, words, words: Mapping the Matter of Climate Change Literature}
\title{A Topography of Climate Change Research}
\author[1,2]{Max Callaghan*}
\author[1,2]{Jan Minx}
\author[2]{Piers M. Forster}

@@

		\noindent We live in an age of ``Big Literature'' 
		\cite{Nunez-Mir2016, Minx2017l}, where the science of climate change is expanding exponentially \cite{Grieneisen2011, Haunschild2016}. In the five years since the publication of the last IPCC assessment report \cite{IPCC2014c}, 202,000 papers were published in the Web of Science (WoS) (see Table \ref{tab}). This is almost as much as the 205,000 papers published during the first five assessment periods; a period of nearly 30 years. A total of around 350,000 new publications can be expected for the current sixth assessment cycle of the IPCC, based on current growth patterns (Figure \ref{pub-growth}. Moreover, the literature has also become more diverse. This is reflected in the expansion of the literature's vocabulary - from 2,000 unique words in the first assessment period to 95,000 words so far in the sixth - indicating that the field has incorporated new content. For example, the zika virus, which was mentioned in 182 articles from 2014-2018, had never before been discussed in the titles or abstracts of articles relating to climate change. Yet it has emerged as a topic of high relevance: the incidence of the virus, the outbreak of which in Brazil in 2016 was declared a public health emergency by the WHO, is set to increase under rising global temperatures \cite{Rao2019}. Similar rapid emergence patterns can be seen for INDCs and SDGs in AR6, and Biochar and REDD in AR5, among others\footnote{The glossary in SI contains a complete list of the acronyms shown in the table}.

		\begin{figure}[htp]
			\begin{center}
				\includegraphics[width=180mm]{plots_pub/pubs_time_wgb.pdf}
				\caption{ The number of climate change documents in the Web of Science in each year. For 2019-21 we project the number of papers assuming there is no more growth, and assuming that growth continues at the same rate as over the past five years}
				\label{pub-growth}
			\end{center}
		\end{figure}

@@

		Big literature poses at least three challenges for scientific policy advice and science itself: \emph{First}, established procedures in scientific assessments like those conducted by the IPCC fail to address the exploding literature base. For example, the ratio of studies cited in IPCC reports to the number or relevant studies has declined from 60\% to 20\%  \cite{Minx2017l}, posing a rapidly growing risk of selection bias. More generally, the provision of comprehensive, objective, open and transparent assessments of the available scientific literature, as defined in the principles governing IPCC work \cite{IPCC2013}, is no longer possible for authors or author teams by traditional means. Machine reading and learning methods as well as other data science applications are required to enable an understanding of the field of climate change research at scale (REFs). \emph{Second}, evidence synthesis - the enterprise of reviewing the literature based on a formal and systematic set of methods \cite{Chalmers2002} - becomes increasingly important for aggregating and consolidating the rapidly emerging knowledge and enabling scientific assessments to do their job. Yet, traditional methods of evidence synthesis themselves are pushed to their limits by the large amount of scientific publications. Whilst the field of evidence synthesis technology that tries to streamline human tasks through machine learning at the different stages of the review process is still in its infancy \cite{Beller2018}. \emph{Finally}, overwhelming amounts of literature may be a major reason why studies of scientific assessments \cite{Bjurström2011, Hulme2010, Victor2015, Corbera2016} do not offer robust quantifications, for claims about the relationship between report citations and the underlying literature.

		This study uses topic modelling \cite{Blei2010} to map out the vast body of evidence on climate change. Topic modelling is an unsupervised machine-learning technique, where patterns of word co-occurrences in documents are used to learn a set of topics which can be used to describe the corpus. The word topic derives from the Greek word for place (topos), and by \textit{situating} the documents in a reduced-form projection of their thematic content (see Figure \ref{oecd_topic_map}), we create a \textit{topographic map} of the literature on climate change. Such a systematic engagement with the thematic content of the climate science is missing from the literature so far. 
		We apply this map in a second step to understand how the IPCC reports have represented the available climate change literature and re-evaluate claims of bias based on a more comprehensive understanding of the available climate science. We enrich the discussion of representation in the literature by discussing topics as well as disciplines.

@@

		Figure \ref{oecd_topic_map} shows a \emph{thematic} or \emph{topopographic map} of the 400,000 publications on climate change in our dataset with a total number of 140 topics. The number of topics must be defined exogenously, but the results are robust to different specifications. Using non-negative matrix factoriziation \cite{Lee1999}, the topics are machine-learned from the papers' abstracts (see methods for details, examples of different model specifications, and a thorough explanation of model selection), and the topic scores of each document are reduced to the two dimensions shown through t-distributed stochastic neighbour embedding \cite{vandermaaten2008}. \footnote{A full list of topics and related words, and a list of documents and related topics are given in the SI}.

@@

		The map shown covers a broad range of topics, with related topics shown in clusters. In general, topics related to climate science and impacts are in the West, while solution-oriented topics are in the East. More fine-grained research areas can also be distinguished. For example, publications related to urban infrastructure (\textbf{buildings}, \textbf{energy}, \textbf{cement}, \textbf{waste}) are located in the East, physical climate impacts such as \textbf{sea-level}, \textbf{droughts}  or [crop] \textbf{yield} are in the South-West and energy systems are in North-East. There are larger groups of documents at the fringes of the map that relate mainly to one or two specific topics such as \textbf{biochar}, \textbf{coral}, or \textbf{CO2 storage}. Interestingly, scenarios feature centrally in the map, at the interface between different scientific communities. This corresponds to their integrative nature in IPCC reports (REF). This map of the thematic structure of the literature could be useful for individual communities or for climate change assessments.

		The disciplinary composition of this research topography indicated by the different colours in Figure \ref{oecd_topic_map} highlights the dominance of natural sciences in climate change research. More than 60\% of the literature is published in natural science journals. Similarly, in 115 out of 140 topics the contribution of publications in natural science journals is greater than any other discipline. We calculate disciplinary entropy of topics as a measure of their degree of interdisciplinarity (see Figure \ref{dis-entropy} and methods for details). This shows how research on health, food, or policy comes from a range of disciplines, while research on ice and oceans comes almost exclusively from the natural sciences).

@@

		Finally, the topography shows the thematic evolution of the literature, with topics exhibiting distinct patterns of growth. Fast-growing topics in the last three assessment periods have included, among others, \textbf{coral}, \textbf{risks}, \textbf{adaptation}, \textbf{hydrogen}, \textbf{buildings}, \textbf{CO2 removal}, \textbf{networks} and \textbf{biochar}. \textbf{Biochar} is particularly remarkable in that the sizeable literature which emerged in AR5 was completely absent from the climate change literature beforehand. 
		The identification of new topics as they emerge, particularly as these are identified without prior knowledge of the literature, can help researchers and assessment-makers to keep abreast of a quickly evolving field.

@@

		Figure \ref{oecd_rep}.a shows that the social sciences were indeed under-represented in the third assessment report, but by the fifth assessment report were over-represented. Likewise, other social sciences than economics have become better represented since AR3  (see figure \ref{subfield}f) with social \& economic geography (4.3\% of the literature), political science (1.0\%), and sociology (0.8\%) showing improved representation in AR5 compared to AR3, and social and economic geography, political science, and other social sciences better represented than economics.

		This challenges what we think we know about the IPCC. The social sciences, by now, are actually the best represented field, with a share in the literature cited by IPCC reports 1.32 times as high as in the literature at large.  On the other hand the Agricultural Sciences and Engineering \& Technology have been consistently under-represented, with 2.27 and 3.49 times the share of studies in the wider literature than in the literature cited by the IPCC in AR5 respectively.

		The topography allows us to delve deeper into the subject matter that receives more or less attention in the IPCC. Figures \ref{oecd_rep}b and \ref{oecd_rep}c plot the representation of the topics shown in the map. Figure \ref{oecd_rep}c shows that topics more commonly cited by IPCC working group I are older and largely better represented in IPCC reports. These topics, for example \textbf{ozone}, \textbf{oceans}, \textbf{clouds}, \textbf{aerosols} and \textbf{sea levels} make up some of the core topics of the physical science of climate change.

		The topics in the lower right of the graph are the most pertinent to the question of whether the IPCC is well representing knowledge on climate change. These topics are newer and until now have been under-represented in IPCC reports. Because they are new areas of knowledge, they may be highly salient in a periodic assessment process. These topics are primarily in working group III, on mitigation \footnote{see methods for a discussion of the categorisation of topics, including CLC, adsorption and hydrogen, which may more properly be described as relevant to WGIII}.

		The difference between these under-represented new topics and other new topics that are better represented is intriguing. This difference is visible in figure \ref{evolution-map}, where in AR5, the clusters of documents around the \textbf{adsorption}, \textbf{buildings}, and \textbf{biochar} topics contain few IPCC citations, whereas the clusters around \textbf{food}, \textbf{health}, \textbf{adaptation}, and \textbf{GHGs} contain more. As shown in figure \ref{oecd_rep}c, \textbf{adsorption}, \textbf{buildings} and \textbf{biochar} are 4.08, 3.34 and 3.61] times more prevalent in the literature than in IPCC citations, while food is 1.22 times more prevalent in the literature and health and adaptation are 1.02 and 2.22 times more prevalent in IPCC citations respectively. The IPCC, has been better at integrating new knowledge from these topics, and in general better at integrating new knowledge from WG II  than WG III topics.

@@

		Recalling policymakers' demands for more solution-oriented assessments \cite{Kowarsch2017}, we could also interpret the topics that are newer and under-represented as ``solutions-relevant''. However, while policymakers' demands for solutions-oriented knowledge were rather about policy options, these under-represented new topics deal with more technical solutions and are found rather in technical disciplines within engineering \& technology and the agricultural sciences.

@@

		\subsection*{Machine-learning for climate change assessments}

		We have shown that social science literature about climate change is over-represented in the IPCC, while technical literature on solutions is under-represented.
		A perfectly proportional representation of every part of the literature is of course not optimal, yet these two features represent new knowledge about the interaction between the IPCC and the literature, and have important implications for the community. 
		The over-representation of the social sciences, combined with the perception that the IPCC needs to include more social science knowledge \cite{Victor2015}, suggests that doing so must involve the funding and production of more social science literature on climate change, not just greater efforts by the IPCC to include it. 
		Moreover, the fact that solutions-relevant topics are under-represented - while policymakers demand more solutions-oriented assessments - and that these topics contain little social science research, suggests areas where social science research may be usefully conducted. Further, this fact can open a debate about the extent to which the IPCC should include more technical knowledge on solutions from disciplines within engineering and the agricultural sciences.

		The map can also serve as a guide to future assessments, to ensure that decisions about how different areas of the literature are represented are well-informed. 
		The map can contribute to these decisions as they are made from the scoping process, through to the selection by authors of individual studies.
		Beyond this, the methods shown here could aid other processes in the production of IPCC reports, such as the identification of potential authors to acheive a better balance across sectors, regions and genders \cite{Corbera2016}.
		Outside of the IPCC, this approach is part of ongoing attempts to make use of machine-learning within evidence synthesis. The topographic map presented is a new approach to rapidly mapping very large literatures.

		The more than 400,000 publications we deal with here represent a wealth of knowledge on climate change and climate solutions. However, we acknowledge that our dataset is by no means exhaustive. We repeat an established query \cite{Haunschild2016}, granting that it may have imperfections. Beyond this we miss publications not in the Web of Science (some small journals, some books, and most grey literature, not to mention indigenous knowledge \cite{Ford2016b}); and studies relevant for the work of the IPCC, but that do not directly mention climate change (for example on energy policy). We argue that this remains a reasonable system boundary given data availabilty, and stress that the documents not included in our study alter our findings only if they have systematically different patterns of citation by the IPCC. In the future, making use of more sources climate change knowledge, and extracting and classifying information from full texts (here we only use abstracts), could improve the granularity of this topography. Most importantly, machine learning applications should be explored that support IPCC authors in assessing the literature. This would prepare IPCC assessments for the age of big literature.

@@

	%\listoffigures
	\linespread{1}
	%\bibliography{Mendeley}

\begin{thebibliography}{10}

@@

	\bibitem{Nunez-Mir2016}
	Gabriela~C. Nunez-Mir, Basil~V. Iannone, Bryan~C. Pijanowski, Ningning Kong,
	Songlin Fei, and Richard Fitzjohn.
	\newblock {Automated content analysis: addressing the big literature challenge
		in ecology and evolution}.
	\newblock {\em Methods in Ecology and Evolution}, 7(11):1262--1272, 2016.

@@

	\bibitem{Lee1999}
	D~D Lee and H~S Seung.
	\newblock {Learning the parts of objects by non-negative matrix factorization.}
	\newblock {\em Nature}, 401(6755):788--91, 1999.

	\bibitem{vandermaaten2008}
	Laurens van~der Maaten and Geoffrey Hinton.
	\newblock {Visualizing Data using t-SNE}.
	\newblock {\em Journal of Machine Learning Research}, 9:2579--2605, 2008.

	\bibitem{Moss2010}
	Richard~H. Moss, Jae~A. Edmonds, Kathy~A. Hibbard, Martin~R. Manning, Steven~K.
	Rose, Detlef~P. {Van Vuuren}, Timothy~R. Carter, Seita Emori, Mikiko Kainuma,
	Tom Kram, Gerald~A. Meehl, John~F.B. Mitchell, Nebojsa Nakicenovic, Keywan
	Riahi, Steven~J. Smith, Ronald~J. Stouffer, Allison~M. Thomson, John~P.
	Weyant, and Thomas~J. Wilbanks.
	\newblock {The next generation of scenarios for climate change research and
		assessment}.
	\newblock {\em Nature}, 463(7282):747--756, 2010.

@@

@@

\end{thebibliography}

	\bibliographystyle{unsrt}

	%	\include{methods}
